From Freebase to Wikidata: The Great Migration

### 2015-10-13 paper.tex
Reworked Sections 4.0 and 4.1.

As outlined earlier, the \emph{Primary Sources Tool} is a~crowdsourced human curation
software solution that displays Freebase statements
for verification to the user that can be added to the currently shown Wikidata item.
The user can, with just one click, reject or approve a~statement,
and, in case of approval, therewith add the statement to Wikidata.
The Primary Sources Tool is currently available as a~so-called gadget,
\emph{i.e.}, as an opt-in feature in Wikidata.
This tool has no dependencies on Freebase and can be easily reused for other projects
that want to import datasets into Wikidata.
Therefore, we have implemented an upload API in the back-end
and a~filter in the front-end that allows users to display only statements from a~given dataset.
This flexibility is much-wanted, and during the development period of the project
another researcher approached us in order to upload a~dataset extracted from natural language text.
To visualize the migration progress, we have created a~realtime dashboard
available at \url{https://tools.wmflabs.org/wikidata-primary-sources/status.html}.
The code of the Primary Sources Tool is openly available
(\url{https://github.com/google/primarysources}) under the terms of the Apache~2.0 license.
At the time of writing (October, 2015), the tool has been used by more than hundred users
who performed about 30,000 approval or rejection actions.
More than 12M~statements have been uploaded in total.
\autoref{fig:primary-sources-tool} shows a~screenshot of the user interface
of the tool that integrates nicely into the Wikidata user interface.

@@

\todo{Denny: write me}

>>>>

As outlined earlier, the \emph{Primary Sources Tool} is a~crowdsourced human curation
software solution that displays Freebase statements for verification to the user
so that these statements can be added to the currently shown Wikidata item.
With just one click, the user can reject or approve a~statement,
and, in case of approval, therewith add the statement to Wikidata.
The code of the Primary Sources Tool is openly available
under the terms of the Apache~2.0 license (\url{https://github.com/google/primarysources}).
The tool is deployed as a~so-called gadget,
\emph{i.e.}, as an easily activatable add-on feature in Wikidata.
It is independent from Freebase and can easily be reused to import other datasets into Wikidata.
Therefore, we have implemented an upload API in the back-end
and a~dataset filter in the front-end that allows users
to display only statements from a~selected dataset.
This flexibility is indeed much-wanted, and during the development period of the project
another researcher approached us in order to upload a~dataset extracted from natural language text.
To visualize the migration progress, we have created a~realtime dashboard
available publicly at \url{https://tools.wmflabs.org/wikidata-primary-sources/status.html}.
At the time of writing (October, 2015), the tool has been used by more than a~hundred users
who performed about 30,000 approval or rejection actions.
More than 12M~statements have been uploaded in total.
\autoref{fig:primary-sources-tool} shows a~screenshot of the user interface
of the tool that integrates nicely with the Wikidata user interface.

@@

In order to prepare the data for integration into the Primary Sources Tool,
we have created a set of scripts which map the content
of the last Freebase dump to Wikidata statements.%
\footnote{The source code is available under the terms of the Apache~2.0 license:
\url{https://github.com/google/freebase-wikidata-converter}}
The Primary Sources Tool was developed in parallel with the mapping tool.
We first created a~small dataset of around 1M statements
based on few select properties (\emph{e.g.}, \texttt{/people/person/birth\_place})
and deployed both the first dataset and a~basic version of the tool
in order to gather initial user feedback.
In the following, we progressively added more statements to the tool's database back-end
and made the opt-in process to use the tool in Wikidata easier for users.
Initially, in order to use the tool, users had to manually copy and paste a~line
in their individual \texttt{common.js} user profiles,
whereas after the conversion to a~Wikidata gadget,
the tool could simply be activated via an option in the user preferences.
Thereby, we were able to both slowly increase the number of users of the tool
and upscale the tool's database back-end without risking performance bottlenecks.
The process of adding data in small batches of 1M~to~2M statements per batch allowed us
to detect some potential performances issues in the tool and fix them
before they became actual problems.
Another positive side-effect of this approach was that it allowed us to gather quick feedback
from the community about the already mapped data and thus to adapt the mapping scripts easily
and correct minor issues with the data early on already in the back-end.
